MATH 245 Homework 5

### 2024-04-09 PDEs/math245_hw5.Rmd
pde hw5 prob 1

## Done: 2, 4, 5, 6
## Not started: 1, 3

@@

@@

The solution of (\ref{eqn:inhomowave}) is then
$$u(t,x)= a_0(t)+\sum_{n=1}^{\infty}  a_n(t) \cos{  \left(\frac{n\pi x}{L}\right) }$$
where $$a_0(t)=\frac{1}{L} \int_{0}^{L} \left[t\,\Psi(x)+\Phi(x) \right]dx+\int_0^t(t-s)q_n(s)\,ds$$

$$a_n(t)=A_n\cos{\left(\frac{n\pi ct}{L}\right)}+B_n\sin{\left(\frac{n\pi ct}{L}\right)}+\int_0^t\frac{\sin{\left(\frac{n\pi c}{L} (t-s)\right)}q_n(s)}{\frac{n\pi c}{L}}ds$$
$$A_n= \frac{2}{L} \int_{0}^{L}\Phi(x)\cos{  \left(\frac{n\pi x}{L}\right) }dx, \quad B_n=\frac{2}{n\pi c}\int_{0}^{L}\Psi(x)\cos{  \left(\frac{n\pi x}{L}\right) }dx$$
$$q_n (t)= \frac{2}{L} \int_{0}^{L}H(t,x)\cos{  \left(\frac{n\pi x}{L}\right) }dx$$
\begin{align*}
H(t,x)&=F(t,x)-w_{tt}+c^2w_{xx} \\
&=F(t,x)-\left[\left(x-\frac{x^2}{2L}\right)h''(t)+\frac{x^2}{2L}g''(t)\right]
+\frac{c^2}{L} \Bigl( g(t)- h(t) \Bigr) \\
\Phi(x)  &= u(0,x)-w(0,x) \\
&= \phi(x) -w(0,x)  \\
&= \phi(x) -\left(x-\frac{x^2}{2L}\right)h(0)-\frac{x^2}{2L}g(0) \\
\Psi(x) &=u_t(0,x)-w_t(0,x) \\
&= \psi(x)  -  w_t(0,x)  \\
&= \psi(x)- \left(x-\frac{x^2}{2L}\right)h'(0)-\frac{x^2}{2L}g'(0)
\end{align*}

@@

Therefore, our solution to (\ref{eqn:beam}) is:
\begin{tcolorbox}[colback=white, title=Solution]
\begin{align*}
& u(t,x)= \sum_{n=1}^{\infty} \left[Ae^{\left(\frac{n\pi}{L}\right)^2 ct}+Be^{-\left(\frac{n\pi}{L}\right)^2 ct} \right]\sin{\left( \frac{n\pi x}{L}\right)}, \quad \text{where} \\ 
& B=\frac{1}{L} \int_{0}^{L} \left[\phi(x)- c\left(\frac{n\pi}{L}\right)^2  \psi(x)\right]\sin{  \left(\frac{n\pi x}{L}\right) }dx\\
& A=\frac{1}{L} \int_{0}^{L} \left[\phi(x)+ c\left(\frac{n\pi}{L}\right)^2  \psi(x)\right]\sin{  \left(\frac{n\pi x}{L}\right) }dx
\end{align*}
\end{tcolorbox}

@@

Therefore, our solution to (\ref{eqn:radio}) is:
\begin{tcolorbox}[colback=white, title=Solution]
\begin{align*}
& u(t,x)= b_0(t)+\sum_{n=1}^{\infty} b_n(t) \sin{(nx)}, \quad \text{where} \\ 
& b_n(t)=\int_{0}^{t} e^{n^2(s-t)}q_n\;ds,\\
& q_n=\frac{2A}{\pi} \frac{n(-1)^{n+1} e^{-a\pi}+n}{n^2+a^2}, \\ 
& b_0(t)=\frac{-At}{a \pi} \left(e^{-a\pi}-1\right)+b_0(0), \\
& b_0(0)=\frac{2}{n\pi}(1+(-1)^{n+1})
\end{align*}
\end{tcolorbox}

@@

Therefore, our solution is:
\begin{tcolorbox}[colback=white, title=Solution]
\begin{equation}\label{eqn:teleSol}
u(t,x)= \sum^{\infty}_{n=1} A_n \,e^{\gamma t} \sin{\left( \frac{n\pi x}{l}\right)}+ \\
\sum^{\infty}_{n=1} B_n \,e^{\zeta t} \sin{\left( \frac{n\pi x}{l}\right)},\end{equation}

>>>>

## Done: 1, 2, 4, 5, 6
## Not started: 3

@@

\begin{equation}\label{eqn:inhomoheatV1}
    \begin{cases}
     v_t-kv_{xx} = x \cos{t}-w_t+kw_{xx}                  & 0<x<\pi, \quad t>0 \\
     v(0,x) = u(0,x)-w(0,x) = \sin{(\pi x)}-w(0,x)  & 0<x<\pi\\
     v(t,0)=u(t,0)-w(t,0) = t^2-t^2=0                         & t>0       \\
     v(t,\pi)=u(t,\pi)-w(t,\pi) = 2t-2t=0                     & t>0
    \end{cases}       
\end{equation}

So we want a function $w(t,x)$ such that $w(t,0)=t^2$ and $w(t,\pi)=2t$. The following function satisfies these requirements, and also has the convenient properties that $w_{xx}(t,x)=0$ and $w(0,x)=0$.

$$w(t,x)=t^2+ \frac{2t-t^2}{\pi}x $$
$$w_t(t,x)=2t+ \frac{2-2t}{\pi}x $$

Thus, set $H(t,x)=x \cos{t}-2t-\frac{2-2t}{\pi}x$, and our problem simplifies to
\begin{equation}\label{eqn:inhomoheatV}
    \begin{cases}
     v_t-kv_{xx} = H(t,x) & 0<x<\pi, \quad t>0 \\
     v(0,x) = \sin{(\pi x)}  & 0<x<\pi\\
     v(t,0)=0                & t>0       \\
     v(t,\pi)=0              & t>0
    \end{cases}       
\end{equation}

### Step 2

Assume (\ref{eqn:inhomoheatV}) has a separated solution $v(t,x)=X(x)T(t)$. Based on the Dirichlet BCs, we will assume the solution takes the form

$$v(t,x)= \sum_{n=1}^{\infty}  b_n(t) \sin{(nx)} $$
Where the $\pi$ in the numerator cancels with $L=\pi$ in the denominator of $\sin{  \left(\frac{n\pi x}{L}\right) }$. Now we substitute this into the PDE in (\ref{eqn:inhomoheatV}):

\begin{equation}\label{eqn:step2a}
H(t,x)=\sum_{n=1}^{\infty}\left[b_n''(t) +  kn^2 b_n(t)\right] \sin{(nx))}
\end{equation}

For each fixed $t$, we write $H(t,x)$ as the Fourier sine series
\begin{equation}\label{eqn:step3a}
H(t,x)=\sum_{n=1}^{\infty}q_n(t) \sin{(nx)} \end{equation}

Such that
\begin{align*}
q_n (t)&= \frac{2}{L} \int_{0}^{L}H(t,x)\sin{  \left(\frac{n\pi x}{L}\right) }dx \\
      &= \frac{2}{\pi} \int_{0}^{\pi}H(t,x)\sin{ (nx)}dx
\end{align*}

By (\ref{eqn:step2a}) and (\ref{eqn:step3a}), we get
$$b_n''(t) +  kn^2 b_n(t)=q_n(t) $$

$$\mu(t) = \exp\left(\int_0^t kn^2ds\right)=e^{kn^2t} $$

$$b_n(t)= \frac{1}{\mu (t)} \left[\int_0^t \mu (s) q_n (s)+b_n(0)\right] $$

$$b_n(t)= e^{-kn^2t}\;b_n(0)+ \int_0^t e^{kn^2(s-t)} q_n (s)$$

To find $b_n(0)$, we use the IC from (\ref{eqn:inhomoheatV}):

$$v(0,x)= \sin{(\pi x)}=\sum_{n=1}^{\infty}  b_n(0) \sin{(nx)} $$

\begin{align*}
b_n(0)& =\frac{2}{\pi} \int_{0}^{\pi} \sin{(\pi x)} \cos{(nx)}dx \\ 
      &=\frac{1}{\pi} \int_{0}^{\pi} \sin{(x(\pi+n))} +\sin{(x(\pi-n))}dx \\                   &=\left[\frac{-\cos{(x(\pi+n))}}{\pi(\pi+n)}\right]^\pi_0+\left[\frac{-\cos{(x(\pi-n))}}{\pi(\pi-n)}\right]^\pi_0 \\
      &=\left[\frac{1-\cos{(\pi(\pi+n))}}{\pi(\pi+n)}\right]+\left[\frac{1-\cos{(\pi(\pi-n))}}{\pi(\pi-n)}\right]\\
      &=2\left[\frac{1+\cos(\pi^2)(-1)^{n+1}}{\pi(\pi+n)}\right]
\end{align*}

Where we used $\cos(\alpha+\beta)=\cos\alpha\cos\beta-\sin\alpha\sin\beta$ to find that
$$\cos{(\pi^2+n\pi)}=\cos(\pi^2)\cos(n\pi)-\sin(\pi^2)\sin(n\pi)=\cos(\pi^2)(-1)^n$$
$$\cos{(\pi^2-n\pi)}=\cos(\pi^2)\cos(n\pi)+\sin(\pi^2)\sin(n\pi)=\cos(\pi^2)(-1)^n$$

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
The solution of (\ref{eqn:inhomoheat}) is then
$$u(t,x)= \sum_{n=1}^{\infty} \Bigl[ b_n(t) \sin{(nx)}\Bigr]+ t^2+ \frac{2t-t^2}{\pi}x$$
where 
$$b_n(t)= 2e^{-kn^2t}\;\left[\frac{1+\cos(\pi^2)(-1)^{n+1}}{\pi(\pi+n)}\right]+ \int_0^t e^{kn^2(s-t)} q_n (s)$$
$$q_n (t)= \frac{2}{\pi} \int_{0}^{\pi}H(t,x)\sin{ (nx)}dx, \quad \quad H(t,x)=x \cos{t}-2t-\frac{2-2t}{\pi}x$$
\end{tcolorbox}

@@

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
The solution of (\ref{eqn:inhomowave}) is then
$$u(t,x)= a_0(t)+\sum_{n=1}^{\infty}  a_n(t) \cos{  \left(\frac{n\pi x}{L}\right) }+\left(x-\frac{x^2}{2L}\right)h(t)+\frac{x^2}{2L}g(t)$$
where $$a_0(t)=\frac{1}{L} \int_{0}^{L} \left[t\,\Psi(x)+\Phi(x) \right]dx+\int_0^t(t-s)q_n(s)\,ds$$

$$a_n(t)=A_n\cos{\left(\frac{n\pi ct}{L}\right)}+B_n\sin{\left(\frac{n\pi ct}{L}\right)}+\int_0^t\frac{\sin{\left(\frac{n\pi c}{L} (t-s)\right)}q_n(s)}{\frac{n\pi c}{L}}ds$$
$$A_n= \frac{2}{L} \int_{0}^{L}\Phi(x)\cos{  \left(\frac{n\pi x}{L}\right) }dx, \quad B_n=\frac{2}{n\pi c}\int_{0}^{L}\Psi(x)\cos{  \left(\frac{n\pi x}{L}\right) }dx$$
$$q_n (t)= \frac{2}{L} \int_{0}^{L}H(t,x)\cos{  \left(\frac{n\pi x}{L}\right) }dx$$
\begin{align*}
H(t,x)&=F(t,x)-w_{tt}+c^2w_{xx} \\
&=F(t,x)-\left[\left(x-\frac{x^2}{2L}\right)h''(t)+\frac{x^2}{2L}g''(t)\right]
+\frac{c^2}{L} \Bigl( g(t)- h(t) \Bigr) \\
\Phi(x)  &= u(0,x)-w(0,x) \\
&= \phi(x) -w(0,x)  \\
&= \phi(x) -\left(x-\frac{x^2}{2L}\right)h(0)-\frac{x^2}{2L}g(0) \\
\Psi(x) &=u_t(0,x)-w_t(0,x) \\
&= \psi(x)  -  w_t(0,x)  \\
&= \psi(x)- \left(x-\frac{x^2}{2L}\right)h'(0)-\frac{x^2}{2L}g'(0)
\end{align*}
\end{tcolorbox}

@@

Therefore, our solution to (\ref{eqn:beam}) is:
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
\begin{align*}
& u(t,x)= \sum_{n=1}^{\infty} \left[Ae^{\left(\frac{n\pi}{L}\right)^2 ct}+Be^{-\left(\frac{n\pi}{L}\right)^2 ct} \right]\sin{\left( \frac{n\pi x}{L}\right)}, \quad \text{where} \\ 
& B=\frac{1}{L} \int_{0}^{L} \left[\phi(x)- c\left(\frac{n\pi}{L}\right)^2  \psi(x)\right]\sin{  \left(\frac{n\pi x}{L}\right) }dx\\
& A=\frac{1}{L} \int_{0}^{L} \left[\phi(x)+ c\left(\frac{n\pi}{L}\right)^2  \psi(x)\right]\sin{  \left(\frac{n\pi x}{L}\right) }dx
\end{align*}
\end{tcolorbox}

@@

Therefore, our solution to (\ref{eqn:radio}) is:
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
\begin{align*}
& u(t,x)= b_0(t)+\sum_{n=1}^{\infty} b_n(t) \sin{(nx)}, \quad \text{where} \\ 
& b_n(t)=\int_{0}^{t} e^{n^2(s-t)}q_n\;ds,\\
& q_n=\frac{2A}{\pi} \frac{n(-1)^{n+1} e^{-a\pi}+n}{n^2+a^2}, \\ 
& b_0(t)=\frac{-At}{a \pi} \left(e^{-a\pi}-1\right)+b_0(0), \\
& b_0(0)=\frac{2}{n\pi}(1+(-1)^{n+1})
\end{align*}
\end{tcolorbox}

@@

Therefore, our solution to (\ref{eqn:tele}) is:
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
\begin{equation}\label{eqn:teleSol}
u(t,x)= \sum^{\infty}_{n=1} A_n \,e^{\gamma t} \sin{\left( \frac{n\pi x}{l}\right)}+ \\
\sum^{\infty}_{n=1} B_n \,e^{\zeta t} \sin{\left( \frac{n\pi x}{l}\right)},\end{equation}
